---
title: "Impact of food production on the world"
author: "Paweł Drągowski & Katarzyna Bownik & Krzysztof Kłopotowski"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      include = TRUE,
                      eval = TRUE,
                      warning = FALSE)
library(dplyr)
library(ggplot2)
library(plotly)
library(readxl)
library(tidyr)
```

# Aim of the project

In this report, we asked ourselves what impact has food production on various
aspects of environment, such as land use, water use and greenhouse gas
emissions. While visualizing our data, we put special emphasis on comparison
between different food products, such as meats, oats, vegetables and others.

# Overall production of food classes in each continent

```{r, file="production.R"}
figprod
```

From the interactive plot we conclude, that cereals are by far most popular type
of food in the world, rivaled only by rice in Asia; its easy to see why, cereals
are excellent source of carbohydrates and many minerals, while being relatively
easy to farm and grow in intermediate climate. Milk is also very prevalent,
especially in northern hemisphere where cows are common. It may seem as a surprise
that chicken and pig meat is comparably popular, mostly due to the appeal of 
swine in east Asian cuisine.

# How damaging for the environment are different food products?

```{r, file="foods.R"}
figcol
```

## Description

In our comparison amount of pollution was measured in three ways: most basic and
intuitive one is measurement by kilograms, but it isn't very indicative of how
resourceful product is for our organism. For that reason, we also evaluated 
impact of products by kilocalories and proteins, since those are easily the most
important statistics for healthy and nutritional diet.  
Each products impact was weighed in three categories: in grams of eutrophying 
emissions (that is effects of enriching bodies of water in minerals during food
production), in square meters of land and liters of water used for production.

## Conclusions

By analysing our plot we reach several conclusions:  
1. Meat production, especially red meat and fishes emits large portions of 
phosphates into atmosphere. Most of animals which place high on land used are
raised in open-field system, which counts for small percent of total output 
of meat. Only poultries can be considered somewhat eco-friendly, with their
intermediate usage of water and emissions.  
2. Rice, as expected is very water-hungry, over twice more than its counterparts
- wheat and rye.  
3. Potatoes, root vegetables and tofu overally score greatly in all categories.

# Where in the world emissions from food production are largest?

```{r, file="map.R"}
figmap
```

## Reading the plot

On hover, each country displays its percentage share to global emissions (that 
means the sum of those percentages equals 100). Countries which did not exist
before year 2015 were ignored.

## Conclusions

First result which we can note is that USA, Brazil, Indonesia and China in total 
are responsible for around 25 to 30 percent of emissions in the world. This is
pessimistic result, as in those countries reside only around 800 million people, 
around tenth of global population. Although country with the highest emissions 
from food production is China, it has a huge population of 1.4 billion to feed.
Even so, a comparable in population country of India has half as much emissions,
on one hand thanks to its more plant based cuisine, but on the other as an effect
of undernourishment and poverty problems there.

# Bibliography

1. [https://www.fao.org/faostat/en/#data](https://www.fao.org/faostat/en/#data)
2. [https://ourworldindata.org/environmental-impacts-of-food](https://ourworldindata.org/environmental-impacts-of-food)
3. [https://www.science.org/doi/10.1126/science.aaq0216](https://www.science.org/doi/10.1126/science.aaq0216)
